---
title: "Endophyte Erythromycin Diffusion Testing"
author: "Grace Carey"
date: "2023-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(tidyverse)
library(dplyr )
library(ggplot2)
library(reshape2)
library(magrittr)
library(knitr)
library(data.table)
```

## Introduction

The purpose of this document is to describe and visualize the results of the KB disk diffusion testing experiment on the AMR STRIPS 2021 root endophytes. 

## Prep

Read data file into R: 

```{r, chunk0, echo=FALSE}

work <- read_csv("KIRBY_BAUER_DEMO_ONLY.csv", col_types =c(PLATE_ID = "factor", ZONE_mm = "numeric", ANTIBIOTIC = "factor"))

view(work)
```


```{r, chunk1, echo=FALSE}
# colname <- colnames(work)
# kable(colname)
```
Samples were taken from the following columns: 

```{r, chunk1, echo=FALSE}

unique(work$COLUMN) %>% as.matrix(COLUMN, ncol=6) -> samples_plated

kable(samples_plated)
```


 

```{r, chunk7, echo=FALSE}
# # Start by finding mean percent change from control to erm for each treatment. The first set of numbers represents the % changes without any outliers removed. The second shows the same with four outliers removed.
# 
# tapply(percwk1$PERCENT_CHANGE_CON_TO_ERM, percwk1$COLUMN_TREATMENT, mean)
# 
# tapply(percwk$PERCENT_CHANGE_CON_TO_ERM, percwk$COLUMN_TREATMENT, mean)
```
# Below this point all is from the cfu study- edit for disks

## Percent change CFU/ml from control to erm plates- no outliers removed


Inspect the data as you want. Next, generate standard deviation for your column of interest. Then plot the mean cfu/g with error bars. 

```{r, chunk8, echo=FALSE}
pwork.summary1 <- percwk1 %>%
  group_by(COLUMN_TREATMENT) %>%
  summarise(
    sd = sd(PERCENT_CHANGE_CON_TO_ERM, na.rm = TRUE),
    PERCENT_CHANGE_CON_TO_ERM = mean(PERCENT_CHANGE_CON_TO_ERM)
  )
pwork.summary <- percwk %>%
  group_by(COLUMN_TREATMENT) %>%
  summarise(
    sd = sd(PERCENT_CHANGE_CON_TO_ERM, na.rm = TRUE),
    PERCENT_CHANGE_CON_TO_ERM = mean(PERCENT_CHANGE_CON_TO_ERM)
  )
```
This shows the summary of the all-sample percent change data
``` {r, chunk9, echo=FALSE}
kable(pwork.summary1) 
```
Doing a one-way anova looking at the effect of Column treatment on percent change from con to erm, using the full data set (no outliers removed): 

``` {r, chunk090, echo=FALSE}
one.way.perc.all <- aov(PERCENT_CHANGE_CON_TO_ERM ~ COLUMN_TREATMENT, data = percwk)
summary(one.way.perc.all) 
```

Now we visualize the percent changes, by treatment, with error bars. 

First the data with no outliers removed: 

``` {r, chunk10, echo=FALSE}
cfu_column_trt.pwrk1<- ggplot(percwk1, aes(COLUMN_TREATMENT, PERCENT_CHANGE_CON_TO_ERM,  color=COLUMN_TREATMENT)) +
  ggtitle("Percent change, control to erm plates, by column treatment, outliers included ") +
  labs(x = "Column Treatment", y = "Percent change") +
  geom_jitter(position = position_jitter(0.2), color = "gray") + 
  geom_pointrange(aes(ymin = PERCENT_CHANGE_CON_TO_ERM -sd, 
                      ymax = PERCENT_CHANGE_CON_TO_ERM +sd),data = pwork.summary1) 
cfu_column_trt.pwrk1
ggsave("Growth by Column Treatment perc all samples.png")
```

## Percent change CFU/ml from control to erm plates- outliers removed

This shows the summary of the outlier-removed percent change data
``` {r, chunk009, echo=FALSE}
kable(pwork.summary) 
```
Results of one-way ANOVA testing the effect of column treatment on percent change from con to erm using the data set with outliers removed: 

``` {r, chunk0901, echo=FALSE}
one.way.perc.cut <- aov(PERCENT_CHANGE_CON_TO_ERM ~ COLUMN_TREATMENT, data = percwk1)
summary(one.way.perc.cut) 
```
Then visualize the data with outliers removed: 

``` {r, chunk00101, echo=FALSE}
cfu_column_trt.pwrk<- ggplot(percwk, aes(COLUMN_TREATMENT, PERCENT_CHANGE_CON_TO_ERM,  color=COLUMN_TREATMENT)) +
  ggtitle("Percent change, control to erm plates, by column treatment, outliers removed") +
  labs(x = "Column Treatment", y = "Percent change") +
  geom_jitter(position = position_jitter(0.2), color = "gray") + 
  geom_pointrange(aes(ymin = PERCENT_CHANGE_CON_TO_ERM -sd, 
                      ymax = PERCENT_CHANGE_CON_TO_ERM +sd),data = pwork.summary) 
cfu_column_trt.pwrk
ggsave("Growth by Column Treatment perc.png")
```
## Summary



