---
title: "Endophyte Erythromycin Diffusion Testing"
author: "Grace Carey"
date: "2023-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(tidyverse)
library(dplyr )
library(ggplot2)
library(reshape2)
library(magrittr)
library(knitr)
library(data.table)
library(xlsx)
```

## Introduction

The purpose of this document is to describe and visualize the results of the KB disk diffusion testing experiment on the AMR STRIPS 2021 root endophytes. The tables and dataframes collected will be exported into the excel file "KB_SUMMARY_061223.xlsx". Figures will be exported into the working directory as individual files. 

## Prep
```{r, chunkT0, echo=FALSE}

metadata <- data.frame(Sheet = c("metadata", "Samples_with_key", "samples_plated", "cultures_plated", "stdev_zone_bycol", "stdev_zone_bytrt"),
                 Content = c("sheet containing metadata", "Full data set including zone size, column ID, culture info, and column treatment", "Full list of columns represented by cultures plated in this document", "Full list of inhibition zones recorded in this document", "standard deviation of inhibition zone, by column ID", "standard deviation of inhibition zone, by column treatment")
                 )
exportmetadata <-as.data.frame(metadata)
write.xlsx(metadata, file="KB_SUMMARY_061223.xlsx", sheetName="metadata", append=TRUE, row.names=FALSE)

```

Read data file into R: 

```{r, chunk0, echo=FALSE}

workaa <- read_csv("KB_zones_working.csv", col_types =c(PLATE_ID = "factor", ZONE_mm = "numeric", Antibiotic = "factor", Column = "factor", Treatment = "factor"))
colkey <- read_csv("Column_key.csv", col_types =c(Column = "factor", Treatment = "factor"))

work <- inner_join(workaa, colkey, by=c('Column'='Column'))
#glimpse(colkey)
exportwork <-as.data.frame(work)
write.xlsx(exportwork, file="KB_SUMMARY_061223.xlsx", sheetName="Samples_with_key", append=TRUE, row.names=FALSE)
```
The working dataframe contains information about sample ID, column treatments, zone of inhibition size, dates, and other notes.  

```{r, chunk1, echo=FALSE}
colname <- colnames(work)
kable(colname, caption = "Columns in the working dataframe")
```
Samples were taken from the following soil columns: 

```{r, chunk01, echo=FALSE}

unique(work$Column) %>% as.matrix(Column) -> samples_plated

kable(samples_plated, caption = "Column Samples Plated")

write.xlsx(samples_plated, file="KB_SUMMARY_061223.xlsx", sheetName="samples_plated", append=TRUE, row.names=FALSE)
```

These samples were plated using the standard Kirby-Bauer disk diffusion protocol. Disks were impregnated with 15ug erythromycin each. Two disks were placed on each plate, and each isolate was spread on two plates. For example, isolate C30_D1_R2_IS1_1 was plated twice: on plate C30_D1_R2_IS1_1 and on plate C30_D1_R2_IS1_2. Each plate contained two diffusion disk zones- zone A and zone B. So, zones C30_D1_R2_IS1_1A and C30_D1_R2_IS1_1B were the two zones taken from plate C30_D1_R2_IS1_1. 

```{r, chunk3, echo=FALSE}

unique(work$PLATE_ID) %>% as.matrix(Column, ncol=6) -> cultures_plated

head_cultures<- head(cultures_plated, 15)
kable(head_cultures, caption = "Cultures Plated")
write.xlsx(cultures_plated, file="KB_SUMMARY_061223.xlsx", sheetName="cultures_plated", append=TRUE, row.names=FALSE)
```
# Examining the Data 

```{r, chunk2, echo=FALSE}

tapply(work$ZONE_mm, work$Treatment, mean) 
work.summary <- work %>%
  group_by(Column) %>%
  summarise(
    sd = sd(ZONE_mm, na.rm = TRUE),
    ZONE_mm = mean(ZONE_mm)
  )

work.summary2 <- work %>%
   group_by(Treatment) %>%
   summarise(
     sd = sd(ZONE_mm, na.rm = TRUE),
     ZONE_mm = mean(ZONE_mm)
   )
```
This shows the summary of the zone of inhibition size (in millimiters) mean and standard deviation for each soil column. 
``` {r, chunk9, echo=FALSE}
kable(work.summary, caption = "Standard Deviation and Mean for Zone Size (mm) by Column") 

work.summ.df<- as.data.frame(work.summary)
write.xlsx(work.summ.df, file="KB_SUMMARY_061223.xlsx", sheetName="stdev_zone_bycol", append=TRUE, row.names=FALSE)

```
This shows the summary of the zone of inhibition size (in millimiters) mean and standard deviation for each soil column treatment. Treatments were digestate, manure, or mineral solution. 
``` {r, chunk6, echo=FALSE}
kable(work.summary2, caption = "Standard Deviation and Mean for Zone Size (mm) by Column Treatment") 

work.summ.df2<- as.data.frame(work.summary2)
write.xlsx(work.summ.df2, file="KB_SUMMARY_061223.xlsx", sheetName="stdev_zone_bytrt", append=TRUE, row.names=FALSE)
```
Doing a one-way anova looking at the effect of Column ID on zone size: 

``` {r, chunk5, echo=FALSE}
anova.zone.column <- aov(ZONE_mm ~ Column, data = work)
summary(anova.zone.column) 

```
Doing a one-way anova looking at the effect of column treatment on zone size: 

``` {r, chunk7, echo=FALSE}
anova.zone.column.trt <- aov(ZONE_mm ~ Treatment, data = work)
summary(anova.zone.column.trt) 
```

Now we visualize zone sizes, by column ID, with error bars. 


``` {r, chunk11, echo=FALSE}
zone.mm.col<- ggplot(work, aes(Column, ZONE_mm,  color=Column)) +
  ggtitle("Zone size (mm) by Column ID") +
  labs(x = "Column", y = "Zone (mm)") +
  geom_jitter(position = position_jitter(0.2), color = "gray") + 
  geom_pointrange(aes(ymin = ZONE_mm -sd, 
                      ymax = ZONE_mm +sd),data = work.summary) 
zone.mm.col 
#ggsave("Zone_size_mm_by_Column_ID")
```

Now we visualize zone sizes, by treatment, with error bars. 


``` {r, chunk12, echo=FALSE}
zone.mm.trt<- ggplot(work, aes(Treatment, ZONE_mm,  color=Treatment)) +
   ggtitle("Zone size (mm) by Column Treatment- Erythromycin, Full Dataset") +
   labs(x = "Column Treatment", y = "Zone (mm)") +
   geom_jitter(position = position_jitter(0.2), color = "gray") + 
   geom_pointrange(aes(ymin = ZONE_mm -sd, 
                       ymax = ZONE_mm +sd),data = work.summary2) 
zone.mm.trt
ggsave("Zone_size_mm_by_Column_Treatment.png")
```


## Summary

To be continued. Please note that (as of 06.12.23) data and figures are using a demo (incomplete) data set only. No conclusions can be made from current iteration. 

