---
title: "Endophyte Erythromycin Diffusion Testing"
author: "Grace Carey"
date: "2023-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(tidyverse)
library(dplyr )
library(ggplot2)
library(reshape2)
library(magrittr)
library(knitr)
library(data.table)
```

## Introduction

The purpose of this document is to describe and visualize the results of the KB disk diffusion testing experiment on the AMR STRIPS 2021 root endophytes. 

## Prep

Read data file into R: 

```{r, chunk0, echo=FALSE}

work <- read_csv("KIRBY_BAUER_DEMO_ONLY.csv", col_types =c(PLATE_ID = "factor", ZONE_ID = "factor", CULTURE_ID = "factor", ZONE_mm = "numeric", ANTIBIOTIC = "factor", COLUMN = "factor"))
#glimpse(work)
```
Column names: 

```{r, chunk1, echo=FALSE}
colname <- colnames(work)
kable(colname)
```
Samples were taken from the following columns: 

```{r, chunk01, echo=FALSE}

unique(work$COLUMN) %>% as.matrix(COLUMN, ncol=6) -> samples_plated

kable(samples_plated, caption = "Column Samples Plated")
```

These samples were plated using the standard Kirby-Bauer disk diffusion protocol. Disks were impregnated with 15ug erythromycin each. Two disks were placed on each plate, and each isolate was spread on two plates. For example, isolate C30_D1_R2_IS1_1 was plated twice: on plate C30_D1_R2_IS1_1 and on plate C30_D1_R2_IS1_2. Each plate contained two diffusion disk zones- zone A and zone B. So, zones C30_D1_R2_IS1_1A and C30_D1_R2_IS1_1B were the two zones taken from plate C30_D1_R2_IS1_1. 

```{r, chunk3, echo=FALSE}

unique(work$CULTURE_ID) %>% as.matrix(COLUMN, ncol=6) -> cultures_plated

kable(cultures_plated, caption = "Cultures Plated")
```
# Examining the Data 

```{r, chunk2, echo=FALSE}

 
tapply(work$ZONE_mm, work$COLUMN, mean)

#tapply(work$ZONE_mm, work$COLUMM_TREATMENT, mean) For when the column treatment column exists in this document
```

Inspect the data as you want. Next, generate standard deviation for your column of interest. Then plot the mean cfu/g with error bars. 

```{r, chunk8, echo=FALSE}
work.summary <- work %>%
  group_by(COLUMN) %>%
  summarise(
    sd = sd(ZONE_mm, na.rm = TRUE),
    ZONE_mm = mean(ZONE_mm)
  )

# work.summary2 <- work %>%
#   group_by(COLUMN_TREATMENT) %>%
#   summarise(
#     sd = sd(ZONE_mm, na.rm = TRUE),
#     ZONE_mm = mean(ZONE_mm)
#   )

```
This shows the summary of the all-sample percent change data
``` {r, chunk9, echo=FALSE}
kable(work.summary, caption = "Standard Deviation and Mean for Zone Size (mm) by Column") 
```

``` {r, chunk6, echo=FALSE}
# kable(work.summary2, caption = "Standard Deviation and Mean for Zone Size (mm) by Column Treatment") 
```
Doing a one-way anova looking at the effect of Column ID on zone size: 

``` {r, chunk5, echo=FALSE}
anova.zone.column <- aov(ZONE_mm ~ COLUMN, data = work)
summary(anova.zone.column) 
```
Doing a one-way anova looking at the effect of column treatment on zone size: 

``` {r, chunk7, echo=FALSE}
# anova.zone.column.trt <- aov(ZONE_mm ~ COLUMN_TREATMENT, data = work)
# summary(anova.zone.column.trt) 
```

Now we visualize zone sizes, by column ID, with error bars. 


``` {r, chunk11, echo=FALSE}
zone.mm.col<- ggplot(work, aes(COLUMN, ZONE_mm,  color=COLUMN)) +
  ggtitle("Zone size (mm) by Column ID") +
  labs(x = "Column", y = "Zone (mm)") +
  geom_jitter(position = position_jitter(0.2), color = "gray") + 
  geom_pointrange(aes(ymin = ZONE_mm -sd, 
                      ymax = ZONE_mm +sd),data = work.summary) 
zone.mm.col 
#ggsave("Zone_size_mm_by_Column_ID")
```

Now we visualize zone sizes, by treatment, with error bars. 


``` {r, chunk12, echo=FALSE}
# zone.mm.trt<- ggplot(work, aes(COLUMN_TREATMENT, ZONE_mm,  color=COLUMN_TREATMENT)) +
#   ggtitle("Zone size (mm) by Column Treatment") +
#   labs(x = "Column Treatment", y = "Zone (mm)") +
#   geom_jitter(position = position_jitter(0.2), color = "gray") + 
#   geom_pointrange(aes(ymin = ZONE_mm -sd, 
#                       ymax = ZONE_mm +sd),data = work.summary2) 
# zone.mm.trt
# ggsave("Zone_size_mm_by_Column_Treatment")
```


## Summary

To be continued. Please note that (as of 2/10/23) data and figures are using a demo (very incomplete) data set only. No conclusions can be made from current iteration. 

